### 2022-07-24 main/paper1.tex
Add files via upload

\section{Introduction}
\label{S:1}
The first default prediction models appeared forty years ago with the first generation model presented by \citet{altman}. This work led to the so-called Altman Z-score formula which uses accounting data to compute the default probability of a firm in the next two years. However, when used for financial firms, Altman's Z-score formula needs to be used with care  because, as I will discuss in this paper, financial firms have to be treated carefully due to their frequent use of off-balance sheet financing. Twenty years later, a second generation of reduced-form models used econometrical tools such as maximum likelihood, probit, and logit regressions. The major drawback of these models is that they do not provide multi-period forecasts. One innovative recent development is the use of doubly stochastic Poisson intensity model combined with multiple logit regression to account for multi-period default probability estimation. This model has been proposed by \citet{DSW} in \textit{Multi-period corporate default prediction with stochastic covariates}. The main contribution over prior work was to exploit the time-series dynamics of the explanatory covariates in order to estimate the likelihood of default over several future periods. Their model employs firm-specific and macroeconomic data to create a Markov state vector $X_t$ in order to compute independent firm default intensities $\lambda(t)$ and other types of exit intensities $\phi(t)$.
\citet{DSW} is the first model capable of multi-period default probability estimation using time dynamics of covariates $X_t$. Applications of \citet{DSW} works are various. We can find them in credit rating by credit rating agencies, banks who want to calculate the minimal amount of capital to be held and other researches analyzing the link between macroeconomic cycles and firm's default probabilities. Covariates used by \citet{DSW} are firm's distance to default, firm's trailing one-year stock return, three-month Treasury bill rate and trailing one-year return on S\&P500. Estimating their model on US-listed Industrial firms between 1980 and 2004, they find that distance to default and the current state of the economy have a significant impact on default hazard rates.

The two papers the closest to my paper are the work from \citet{DSW} and \citet{Duan2012}. The first model uses the doubly stochastic argument to derive multi-period default probabilities. To do so, it requires some strong assumptions (i.e Vectorial Autoregressive process) regarding the behavior of the time-series of covariates to generate future random values for the covariates. Obviously, if the process is misspecified, biases are introduced both in the forecasted covariates and in the future default probabilities. Five years later, \citet{Duan2012} shows that we can relax the VAR assumption with the use of forward intensities. This paper explains how we can reduce biases by projecting current event realizations on past data. For convenience, they specify the intensities as a linear function of state variables. I wish to extend the latter by removing the assumption of linear intensities and use an artificial neural network to estimate the intensities of the Poisson processes governing both default and other exits. 
Many papers on estimating default probabilities with machine learning techniques have already been written. For instance, \citet{barboza2017machine} test several machine learning models (support vector machines, bagging, boosting, and random forest) to predict bankruptcy one year prior to the event. They report a substantial improvement in prediction accuracy using these machine learning techniques especially when, in addition to the original Altman’s Z-score variables, six complementary financial indicators are included. To the best of my knowledge, such types of models do not rely on theoretical foundations and this is where the forward intensity model can contribute to the literature. It is able to provide multi-period predictions and is supported by a solid mathematical and econometrical background.

In \citet{DSW}, one of the main assumption is that the covariates governing both default and other exits intensities follow a high-dimensional vector auto-regressive (VAR) process. Using this type of process forces the model to greatly reduce either the number of firms in the sample or the number of state variables explaining firm attributes; if one does not restrict the number of firms or variables in the estimation, the dimension of the model will simply be too high and it will considerably increase computational time. A major step forward made in \citet{Duan2012} was to get rid off the VAR process in order to reduce computational time by using ``a new reduced-form approach based on a \textit{forward} intensity model''. These forward intensities produce a term structure of bankruptcy probabilities without using any sort of high-dimensional process. Using this method allows the model to incorporate a lot more state variables or individual firms in the sample. Moreover, \citet{Duan2012} state that their model may also improve robustness to misspecification because in a VAR model, estimation of future values are highly sensitive to any biases. On the other side, the forward intensity model approach is a ``direct projection of past data on current realizations'' which does not involve random estimation of future values. \\
Regarding covariates used, both \citet{Duan2012} and \citet{DSW} estimate their own Distance-to-Default (hereafter DtD). An important aspect to highlight is that distance to defaults specified in \citet{DSW} differ from those estimated in \citet{Duan2012} since the first one are estimated using the Variance Restriction method (see \textit{Measuring Distance-to-Default for Financial and Non-Financial Firms} \citep{Duan2012DTD}) and the second one using the transformed-data maximum likelihood estimation method to account for financial firms. The variance restriction method is a popular way to implement the \citet{Merton1974} model but fails to estimate properly the default point for financial firms. Following the KMV assumption (see \citet{KMV}), the default point in this method is specified as short term debt plus one half of long term debt and does not take into account other liabilities. However, it is well-known that financial firms such as banks specify a high portion of their debt as ``other liabilities". Hence, to include financial firms in the sample, the default point has to be adjusted to the sum of short term debt, one half of long term debt and a fraction $\delta$ of other liabilities. The method used by \citet{Duan2012} is using a maximum likelihood estimation in order to estimate the unknown fraction $\delta$. The Appendix provides additional methodological information on the DtD estimation using the variance restriction method and the maximum likelihood estimation. \\ 
To summarize, the forward intensity models require an assumption to link covariates to intensities. \citet{Duan2012} uses a linear assumption and a maximum likelihood to estimate those parameters. This paper contributes to previous literature by using neural networks to relax the linear assumption of forward intensities. Neural networks allow the estimation of highly non-linear functions without specifying the form of the relationships. The organization of the paper is as follows. Section 2 sets up the reduced-form model of default, develops the likelihood function used as loss function later on and describes the neural network approach. Section 3 discusses summary statistics of the dataset used. Section 4 presents the results. Section 5 concludes. Appendices at the end of the paper contains several proofs and more details on the Distance-to-Default estimation.

\section{Methodology}
\label{S:2}

@@

Let us denote by $Z_{it}$ the set of firm-specific variables at time t for the firm i and $Y_t$ the set of macroeconomic variables at time t. The first time of entry of the firm i is denoted  $t_i^0$. The econometrician's information set $\mathscr{F}_t$ at time t is thus
\begin{equation}
    \mathscr{F}_t = \{ Y_s : s \leq t \} \cup \mathscr{G}_{1t}  \cup \mathscr{G}_{2t} ... \cup \mathscr{G}_{Nt},
\end{equation}
where 
\begin{equation}
    \mathscr{G}_{it}  = \{  (1_{\tau_{C_i} < u} , 1_{\tau_{D_i} < u} , Z_{iu}) : t_i^0 \leq u \leq \min(\tau_D, \tau_C, t)  \}.
\end{equation}

@@

In Figure \ref{fig:lifespan}, each dot corresponds to a period. At each period, the econometrician gathers firm-specific variables (DTD, Cash/TA, NI/TA, ...) and macroeconomic variables (S\&P500 return, treasury rate). For a particular point in time t, the econometrician knows the whole time-serie of macroeconomic variables until t irrespective of $t_i^0$, and the time-series of firm-specific variables from $t_i^0$ to $\tau_{D_i}$ if $t_i^0 \leq t$ and $t \leq \tau_{D_i}$.
Following Proposition 1 in \citet{DSW}, the conditional probability of default within $s$ years can be computed as 
\begin{equation} \label{eq:3.6}
\mathbb{P}[ \tau_D <t+s |\mathscr{F}_t] = E_t \left[ \int_t^{t+s}  e^{-\int_t^z (\lambda(u)+\phi(u)) du} \cdot \lambda(z) dz \right].
\end{equation}
The probability of default is a function of intensities $\lambda$ (default) and $\phi$ (other exits). However, these intensities are unknown and unobservable. In DSW model, the state variables governing Poisson intensities are assumed to follow a specific vector autoregressive (VAR) process. This assumption is relaxed in Duan's paper by using forward intensity rates. Instead of modeling $\lambda_{it}$ and $\phi_{it}$ as some function of state variables available at time t, \citet{Duan2012} propose to deal with $f_{it}(\tau)$ and $g_{it}(\tau)$ directly as functions of state variables available at time t and the forward starting time of interest $\tau$. The analogy to interest rates would be that $\lambda_t$ is the short rate and $f_t(u)$ is the forward rate for horizon $u$. \citet{Duan2012} propose a model to predict corporate defaults at multiple horizons by estimating these forward intensities via maximum likelihood. To do so, they use a linear assumption in the relationship between the variables and the forward intensities (i.e $f_{it}(\tau) = \exp(\alpha_0(\tau) + \alpha_1(\tau) x_{it,1} + \alpha_2(\tau) x_{it,2} + ... + \alpha_k(\tau) x_{it,k}$). However, it is highly likely that default intensities depend on those covariates in a non-linear way. I propose to use an artificial neural network (ANN) to find the set of weights governing the process $f_{it}$ and $g_{it}$. I show that I am able to capture potential non-linear relationships between the state variables and the forward intensities which significantly improves forecasts.

@@

Since we do not have the exact knowledge of $\lambda$ and $\phi$, \citet{Duan2012} propose to use forward intensity rates. However, we need to translate the default probability given in equation \ref{eq:3.6} (which depends on spot intensities) into a formula that depends on forward intensities. To do so, we first compute the probability of surviving as a function of forward combined intensity, which will be later used to compute the probability of default as a function of both combined and default forward intensities. Let us denote by $F_{it}(\tau)$ the conditional distribution function of the combined (default and other exits) exit time evaluated at $t+\tau$. Hence, 1-$F_{it}(\tau)$ is the probability of surviving in the interval $[t, t+\tau ]$. Therefore, we have :
\begin{equation}
1-F_{it}(\tau) = \mathbb{E}[e^{-\int_t^{t+\tau} (\lambda(s)+\phi(s)) ds}].
\end{equation}
Then, let us introduce the quantity $\psi_{it}(\tau)$ to be :
\begin{equation} \label{eq:3.8}
\psi_{it}(\tau) \equiv -\frac{\ln(1-F_{it}(\tau))}{\tau} \equiv -\frac{\ln(\mathbb{E}[e^{-\int_t^{t+\tau} (\lambda(s)+\phi(s)) ds}])}{\tau}.
\end{equation}
Reverting equation \ref{eq:3.8} gives :
\begin{equation}\label{eq:psi}
e^{-\psi_{it}(\tau)\cdot\tau} = 1-F_{it}(\tau).
\end{equation}
Where $e^{-\psi_{it}(\tau)\cdot\tau}$ is again the survival probability. We now need to compute $\psi_{it}(\tau)\cdot\tau$. At this point, \citet{Duan2012} makes the assumption that $\psi_{it}$ is differentiable and defines the forward combined exit intensity as 
\begin{equation} \label{eq:3.10}
g_{it}(\tau) \equiv \frac{F_{it}'}{1-F_{it}}.
\end{equation}
Equation \ref{eq:3.10} comes from the definition of a hazard rate function. Referring  P.87 of \citet{book}, the definition of a hazard rate function is the following:
%\newtheorem*{mydef}{Definition}
\begin{definition}[Hazard rate]
Let denote by $\tau$ a stopping time and $F(T) \equiv \mathbb{P}[\tau \leq T]$ its distribution function. Assume that $F(T) < 1 \forall T$, and that F(T) has a density f(T). The hazard rate function h of $\tau$ is :
\begin{equation*}
h(T) \equiv \frac{f(T)}{1-F(T)}.
\end{equation*}
\end{definition}
A hazard rate is the local arrival probability of a stopping time per time interval. Please note that under suitable regularity conditions, intensities and hazard rates are closely similar. In particular, in our doubly-stochastic framework, hazard rates and intensities are equivalent. This is why in \citet{Duan2012}, we have $\lambda(t) = h(t)$ and the distinction between hazard rates and intensity is often not made. \\

@@

\noindent Then, we can compute the quantity $\psi_{it}(\tau) \tau$ that we were looking for as :
\begin{equation}
    \psi_{it}(\tau) \tau = \int_0^{\tau}g_{it}(s) ds.
\end{equation}

\noindent For an interested reader, proofs of the above formulations can be found in the Appendix. Hence, the probability of surviving over [t,t+$\tau$] is given by
\begin{equation} \label{survprob}
    \mathbb{P}[\tau_c > t+\tau |\mathbb{F}_t] = \exp(-\int_0^\tau g_{it}(s) ds).
\end{equation}

The forward default intensity for horizon $\tau$ is defined as the limit for a small time step of the probability of defaulting in this small time step given that the firm has survived. The probability is Bayesian and the forward default intensity denoted $f_{it}(\tau)$ is the following :
\begin{equation}
f_{it}(\tau) \equiv \dfrac{\lim_{\Delta t\to 0} \dfrac{\mathbb{P}[t + \tau < \tau_{Di}=\tau_{Ci} \leq t+ \tau + \Delta t]}{\Delta t}}{e^{-\psi_{it}(\tau)\cdot\tau}}.
\end{equation}
Hence, the probability of defaulting between t and t+$\tau$ is given by :
\begin{equation}\label{probdef}
\int_0^{\tau} e^{-\psi_{it}(s)s} f_{it}(s)ds = \int_0^{\tau} e^{-\int_0^s g_{it}(u) du} f_{it}(s)ds.
\end{equation}

@@

In this setup, the likelihood function depends on three types of probabilities (default, other exit and surviving) which themselves depend on the types of intensities (default and other exits). The negative log-likelihood function has to be adjusted to the neural network framework and can be used as objective function to be minimized as long as we feed mini-batches and not single data points (``online learning''). Mini-batch feeding is a common practice in machine learning papers and consists of splitting the available data in batches of fixed size. Then, each backward pass takes one batch to perform one gradient decent step.  \\

To allow further comparison with \citet{Duan2012},  I employ the same discretization of time: $t = 0, 1, 2,...$ and $\tau = 0, 1, 2,...$ are time sequences of one month increment. Similarly, $f_{it}(\tau)$ and $g_{it}(\tau)$ are forward intensities computed at time t for the period [t+$\tau$, t+$\tau$+1]. The use of the $\tau$ index is to account for multiperiod prediction. When $\tau = 0$, the forward intensity model is computing spot intensities. When we set $\tau = 1$, the forward intensity model produces estimates one step ahead, and so on so forth. I denote $X_{it} = (x_{it,1}, x_{it,2} ,...)$ the set of firm-specific and macroeconomic variables explaining both default and combined exit intensities. As specified in \citet{Duan2012} $f_{it}(\tau)$ and $g_{it}(\tau)$ are functions of $X_{it}$ and can be specified as any form of function as long as they satisfy the following constraints. Since combined exit intensity has to be greater or equal than default intensity, we need to make sure that the forms specified for $f_{it}(\tau)$ and $g_{it}(\tau)$ satisfy the following conditions : $f_{it}(\tau) \leq g_{it}(\tau)$, $f_{it}(\tau) > 0$, $g_{it}(\tau) > 0$. \\

I have designed two neural networks, one will be trained to compute $f_{it}$ and one will be trained to output $h_{it}$ where $g_{it} = f_{it} + h_{it}$. I impose non-negativity on outputs of both models such that the combined exit intensity will never be smaller than the default intensity for all horizons. Let us denote by $\lambda$ and $\mu$ the set of parameters (weights) tuned in the neural network for $f_{it}$ and $h_{it}$ respectively. $N^{(\lambda)}$ and $N^{(\mu)}$ stand for the output of the neural network for $f_{it}$ and $h_{it}$ respectively. The log-likelihood for prediction horizon $\tau$ is expressed as
\begin{equation} \label{decomp1}
\mathcal{L}(\lambda(s)) = \sum_{i=1}^N \sum_{t=0}^{T-s-1} \mathcal{L}_{i,t}(\lambda(s)), \qquad s = 0,1,...,\tau-1
\end{equation}

@@

Proof of the above formulation can be found in the Appendix. As in previous studies, the likelihood functions still exhibit the decomposable property which allows to estimate the model for each horizon of prediction independently.

Since the intensities are directly driven by the covariates, \citet{Duan2012} requires an assumption on the mapping from the covariates to these intensities. In \citet{Duan2012} the mapping is made with a linear assumption, whereas in the framework of this paper, the mapping depends on the whole architecture of the neural network. When the neural network has only one hidden layer of one neuron coupled with an exponential activation function, the model boils down to \citet{Duan2012} as the intensities will be a linear combination of the covariates. As the width and depth of the network increases, we depart more and more from the linear assumption and we allow more non-linearities to be incorporated in these intensities. \\

@@

Neural networks can be seen as a very general function to map a given input (in this case a set of macroeconomic and firm-specific variables) into a desired output (forward intensities). It learns how to compute the output by tuning weights in order to minimize a given loss function. A neural network is constructed by juxtaposing several ``hidden'' layers of neurons. The input of each layer is a data transformation of the output of the previous layer. Initially, the weights of the network are assigned random values. Then, the training process starts and consists of many iterations of a forward pass and a backward pass. The forward pass takes as input a batch of data and computes the loss value, the backward pass then computes the gradient and adjust the weights of the network based on a learning rate hyperparameter. As an illustration, figure \ref{fig:net} shows how a neural network looks like with 2 hidden layers : 5 neurons in the first layer and 3 neuron in the second layer. In this example, 3 features (inputs) are fed to the network. Each feature is connected to the first hidden layer by a set of weights. The outputs of the first layer are also weighted to produce the inputs of the second hidden layer. Non-linearity is introduced in each node with a non-linear activation function (i.e sigmoid). Finally, the output of the second hidden layer is aggregated to produce the final output of the model. The neural networks in this paper are implemented in Python using the library TensorFlow. Approximately one hour of computing time is needed to fit all networks on a 32GB RAM quad core 2.7 GhZ computer, and GPU computing is not necessary as the networks are rather small.

@@

    % Annotate the layers
    \node[annot,left of=hl1] {Input Layer};
    \node[annot,right of=hl\Nhidden] {Output Layer};
\end{tikzpicture}
    \mycaption{Illustration of a neural network \text{[}5, 3\text{]}}{In this network, there are three input features, five neurons in the first hidden layer and 3 neurons in the second hidden layer. Each neuron is activated with a sigmoid function.}
    \label{fig:net}
\end{figure}

>>>>

\section{Introduction}
\label{S:1.1}
The first default prediction models appeared forty years ago with the first generation model presented by \citet{altman}. This work led to the so-called Altman Z-score formula which uses accounting data to compute the default probability of a firm in the next two years. However, when used for financial firms, Altman's Z-score formula needs to be used with care  because, as I will discuss in this chapter, financial firms have to be treated carefully due to their frequent use of off-balance sheet financing. Twenty years later, a second generation of reduced-form models used econometrical tools such as maximum likelihood, probit, and logit regressions. The major drawback of these models is that they do not provide multi-period forecasts. One innovative recent development is the use of doubly stochastic Poisson intensity model combined with multiple logit regression to account for multi-period default probability estimation. This model is proposed by \citet{DSW} in \textit{Multi-period corporate default prediction with stochastic covariates}. The main contribution over prior work is to exploit the time-series dynamics of the explanatory covariates in order to estimate the likelihood of default over several future periods. Their model employs firm-specific and macroeconomic data to create a Markov state vector $X_t$ in order to compute independent firm default intensities $\lambda(t)$ and other types of exit intensities $\phi(t)$.
The model proposed in \citet{DSW} is the first one capable of multi-period default probability estimation using time dynamics of covariates $X_t$. Applications of \citet{DSW} works are various. We can find them in credit rating by credit rating agencies, banks who want to calculate the minimal amount of capital to be held and other researches analyzing the link between macroeconomic cycles and firm's default probabilities. Covariates used by \citet{DSW} are firm's Distance-to-Default, firm's trailing one-year stock return, three-month Treasury bill rate and trailing one-year return on S\&P500. Estimating their model on US-listed industrial firms between 1980 and 2004, they find that Distance-to-Default and the current state of the economy have a significant impact on default hazard rates.

\clearpage
The two papers the closest to this chapter are the work from \citet{DSW} and \citet{Duan2012}. The first model uses the doubly stochastic argument to derive multi-period default probabilities. To do so, it requires strong assumptions (e.g. vector autoregressive process) regarding the behavior of the time-series of covariates to generate future random values for the covariates. If the process is misspecified, biases are introduced both in the forecasted covariates and in the future default probabilities. Five years later, \citet{Duan2012} show that we can relax the VAR assumption with the use of forward intensities. This chapter explains how we can reduce biases by projecting current event realizations on past data. For convenience, they specify the intensities as a linear function of state variables. I wish to extend the latter by removing the assumption of linear intensities and use an artificial neural network to estimate the intensities of the Poisson processes governing both default and other exits. 
Machine learning techniques for default probabilities estimation are increasingly drawing attention. \citet{barboza2017machine} test several machine learning models (support vector machines, bagging, boosting, and random forest) to predict bankruptcy one year prior to the event. They report a substantial improvement in prediction accuracy using these machine learning techniques especially when, in addition to the original Altman’s Z-score variables, six complementary financial indicators are included. However, pure data-driven models often lacks economic relevance and this is where the forward intensity model can contribute to the literature. It is able to provide multi-period predictions while being supported by an economic and econometric background.

In \citet{DSW}, one of the main assumption is that the covariates governing both default and other exits intensities follow a high-dimensional vector autoregressive (VAR) process. Using this type of process forces the model to greatly reduce either the number of firms in the sample or the number of state variables explaining firm attributes; if one does not restrict the number of firms or variables in the estimation, the dimension of the model will simply be too high and it will considerably increase computational time. A major step forward made in \citet{Duan2012} was to get rid off the VAR process in order to reduce computational time by using a new reduced-form approach based on a forward intensity model'. These forward intensities produce a term structure of bankruptcy probabilities without using any sort of high-dimensional process. Using this method allows the model to incorporate a lot more state variables or individual firms in the sample. Moreover, \citet{Duan2012} state that their model may also improve robustness to misspecification because in a VAR model, estimation of future values are highly sensitive to any bias. On the other side, the forward intensity model approach is a projection of past observations on current realizations, which does not involve random estimation of future values.

Regarding covariates used, both \citet{Duan2012} and \citet{DSW} estimate their own Distance-to-Default (hereafter DtD). An important aspect to highlight is that Distances-to-Default specified in \citet{DSW} differ from those estimated in \citet{Duan2012} since the former estimates DtD using the variance restriction method (see \citep{Duan2012DTD}) while the latter uses the transformed-data maximum likelihood estimation method to account for financial firms. The variance restriction method is a popular way to implement the \citet{Merton1974} model but fails at estimating properly the default point for financial firms. Following the KMV assumption (see \citet{KMV}), the default point in this method is specified as short-term debt plus one half of long-term debt and does not take into account other liabilities. However, it is well-known that financial firms such as banks specify a high portion of their debt as other liabilities. Hence, to include financial firms in the sample, the default point has to be adjusted to the sum of short-term debt, one half of long-term debt and a fraction $\delta$ of other liabilities. The method used by \citet{Duan2012} is using a maximum likelihood estimation in order to estimate the unknown fraction $\delta$. The Appendix provides additional methodological information on the DtD estimation using the variance restriction method and the maximum likelihood estimation.

To summarize, the forward intensity model requires an assumption to link covariates to intensities. \citet{Duan2012} use a linear assumption and a maximum likelihood to estimate those parameters. This chapter contributes to previous literature by using neural networks to relax the linear assumption of forward intensities. Neural networks allow the estimation of highly non-linear functions without specifying the form of the relationships. The remainder of this chapter is structured as follows. Section \ref{S:1.2} sets up the reduced-form model of default, develops the likelihood function used as loss function later on and describes the neural network approach. Section \ref{S:1.3} discusses summary statistics of the dataset used. Section \ref{sec4} presents the results. Section \ref{S.1.5} concludes. Appendix \ref{appendix1} at the end of the thesis contains several proofs and more details on the Distance-to-Default estimation.

\section{Methodology}
\label{S:1.2}

@@

Let us denote by $Z_{it}$ the set of firm-specific variables at time t for the firm i and $Y_t$ the set of macroeconomic variables at time t. Let  $t_i^0$ be the first time of entry of firm $i$ in the dataset. The econometrician's information set $\mathscr{F}_t$ at time t is thus
\begin{equation}
    \mathscr{F}_t = \{ Y_s : s \leq t \} \cup \mathscr{G}_{1t}  \cup \mathscr{G}_{2t} ... \cup \mathscr{G}_{Nt},
\end{equation}
where 
\begin{equation}
    \mathscr{G}_{it}  = \{  (1_{\tau_{C_i} < u} , 1_{\tau_{D_i} < u} , Z_{iu}) : t_i^0 \leq u \leq \min(\tau_D, \tau_C, t)  \}.
\end{equation}

@@

In Figure \ref{fig:lifespan}, each dot corresponds to a period. At each period, the econometrician gathers firm-specific variables (DtD, Cash/Total Assets, Net Income/Total Assets, ...) and macroeconomic variables (S\&P500 return, treasury rate). For a particular point in time $t$, the econometrician knows the whole time-serie of macroeconomic variables until $t$ irrespective of $t_i^0$, and the time-series of firm-specific variables from $t_i^0$ to $\tau_{D_i}$ if $t_i^0 \leq t$ and $t \leq \tau_{D_i}$.
Following Proposition 1 in \citet{DSW}, the conditional probability of default within $s$ years can be computed as 
\begin{equation} \label{eq:3.6}
\mathbb{P}[ \tau_D <t+s |\mathscr{F}_t] = E_t \left[ \int_t^{t+s}  e^{-\int_t^z (\lambda(u)+\phi(u)) du} \cdot \lambda(z) dz \right].
\end{equation}
The probability of default is a function of intensities $\lambda$ (default) and $\phi$ (other exits). However, these intensities are unknown and unobservable. In DSW model, the state variables governing Poisson intensities are assumed to follow a specific vector autoregressive (VAR) process. This assumption is relaxed in Duan's paper by using forward intensity rates. Instead of modeling $\lambda_{it}$ and $\phi_{it}$ as some function of state variables available at time t, \citet{Duan2012} propose to deal with $f_{it}(\tau)$ and $g_{it}(\tau)$ directly as functions of state variables available at time t and the forward starting time of interest $\tau$. The analogy to interest rates would be that $\lambda_t$ is the short rate and $f_t(u)$ is the forward rate for horizon $u$. \citet{Duan2012} propose a model to predict corporate defaults at multiple horizons by estimating these forward intensities via maximum likelihood. To do so, they use a linear assumption in the relationship between the variables and the forward intensities (i.e. $f_{it}(\tau) = \exp(\alpha_0(\tau) + \alpha_1(\tau) x_{it,1} + \alpha_2(\tau) x_{it,2} + ... + \alpha_k(\tau) x_{it,k}$). However, it is highly likely that default intensities depend on those covariates in a non-linear way. I propose to use an artificial neural network (ANN) to find the set of weights governing the process $f_{it}$ and $g_{it}$. I show that I am able to capture potential non-linear relationships between the state variables and the forward intensities which significantly improves forecasts.

@@

Since we do not have the exact knowledge of $\lambda$ and $\phi$, \citet{Duan2012} propose to use forward intensity rates. However, the default probability given in equation \ref{eq:3.6} (which depends on spot intensities) needs to be translated into a formula that depends on forward intensities. To do so, we first compute the survival probability as a function of forward combined intensity, which will be later used to compute the probability of default as a function of both combined and default forward intensities. Let us denote by $F_{it}(\tau)$ the conditional distribution function of the combined (default and other exits) exit time evaluated at $t+\tau$. Hence, 1-$F_{it}(\tau)$ is the probability of surviving in the interval $[t, t+\tau ]$. Therefore, we have :
\begin{equation}
1-F_{it}(\tau) = \mathbb{E}[e^{-\int_t^{t+\tau} (\lambda(s)+\phi(s)) ds}].
\end{equation}
Next, let us introduce the quantity $\psi_{it}(\tau)$ to be :
\begin{equation} \label{eq:3.8}
\psi_{it}(\tau) \equiv -\frac{\ln(1-F_{it}(\tau))}{\tau} \equiv -\frac{\ln(\mathbb{E}[e^{-\int_t^{t+\tau} (\lambda(s)+\phi(s)) ds}])}{\tau}.
\end{equation}
Reverting equation \ref{eq:3.8} gives :
\begin{equation}\label{eq:psi}
e^{-\psi_{it}(\tau)\cdot\tau} = 1-F_{it}(\tau).
\end{equation}
Where $e^{-\psi_{it}(\tau)\cdot\tau}$ is again the survival probability. We now need to compute $\psi_{it}(\tau)\cdot\tau$. At this point, \citet{Duan2012} make the assumption that $\psi_{it}$ is differentiable and define the forward combined exit intensity as 
\begin{equation} \label{eq:3.10}
g_{it}(\tau) \equiv \frac{F_{it}'}{1-F_{it}}.
\end{equation}
Equation \ref{eq:3.10} comes from the definition of a hazard rate function. Referring \citet{book}, the definition of a hazard rate function is the following:
\begin{definition}[Hazard rate]
Let $\tau$ be a stopping time and $F(T) \equiv \mathbb{P}[\tau \leq T]$ its distribution function. Assume that $F(T) < 1 \forall T$, and that F(T) has a density f(T). The hazard rate function $h$ of $\tau$ is :
\begin{equation*}
h(T) \equiv \frac{f(T)}{1-F(T)}.
\end{equation*}
\end{definition}
A hazard rate is the local arrival probability of a stopping time per time interval. Under suitable regularity conditions, intensities and hazard rates are closely similar. In particular, in our doubly-stochastic framework, hazard rates and intensities are equivalent. Hence, in \citet{Duan2012} and this chapter, $\lambda(t) = h(t)$ and the distinction between hazard rates and intensity is not made. \\

@@

\noindent Next, we can compute\footnote{Proofs of the following formulations can be found in the Appendix \ref{appendix1}.} the quantity $\psi_{it}(\tau) \tau$ that we were looking for as :
\begin{equation}
    \psi_{it}(\tau) \tau = \int_0^{\tau}g_{it}(s) ds.
\end{equation}

\noindent Hence, the probability of surviving over [t,t+$\tau$] is given by :
\begin{equation} \label{survprob}
    \mathbb{P}[\tau_c > t+\tau |\mathbb{F}_t] = \exp(-\int_0^\tau g_{it}(s) ds).
\end{equation}

The forward default intensity for horizon $\tau$ is defined as the limit for a small time step of the probability of defaulting in this small time step given that the firm survives until the considered horizon. The probability is Bayesian and the forward default intensity denoted $f_{it}(\tau)$ is the following :

\begin{equation}
f_{it}(\tau) \equiv \dfrac{\lim_{\Delta t\to 0} \dfrac{\mathbb{P}[t + \tau < \tau_{Di}=\tau_{Ci} \leq t+ \tau + \Delta t]}{\Delta t}}{e^{-\psi_{it}(\tau)\cdot\tau}}.
\end{equation}

Hence, the probability of defaulting between t and t+$\tau$ is given by :

\begin{equation}\label{probdef}
\int_0^{\tau} e^{-\psi_{it}(s)s} f_{it}(s)ds = \int_0^{\tau} e^{-\int_0^s g_{it}(u) du} f_{it}(s)ds.
\end{equation}

@@

In this setup, the likelihood function depends on three types of probabilities (default, other exit and surviving) which themselves depend on two types of intensities (default and other exits). The negative log-likelihood function has to be adjusted to the neural network framework and can be used as objective function to be minimized as long as we feed mini-batches and not single data points (``online learning''). Mini-batch feeding is a common practice in machine learning and consists of splitting the available data in subsets of fixed size. Next, each backward pass takes one batch to perform gradient descent to update the parameters of the model. \\

To allow further comparison with \citet{Duan2012},  I employ the same discretization of time: $t = 0, 1, 2,...$ and $\tau = 0, 1, 2,...$ are time sequences of one month increment. Similarly, $f_{it}(\tau)$ and $g_{it}(\tau)$ are forward intensities computed at time t for the period [t+$\tau$, t+$\tau$+1]. The use of the $\tau$ index is to account for multiperiod prediction. When $\tau = 0$, the forward intensity model computes spot intensities. When we set $\tau = 1$, the forward intensity model produces estimates one step ahead, and so on so forth. I denote $X_{it} = (x_{it,1}, x_{it,2} ,...)$ the set of firm-specific and macroeconomic variables explaining both default and combined exit intensities. As specified in \citet{Duan2012}, $f_{it}(\tau)$ and $g_{it}(\tau)$ are functions of $X_{it}$ and can be specified as any form of function as long as they satisfy the following constraints. Since combined exit intensity has to be greater or equal than default intensity, we need to make sure that the forms specified for $f_{it}(\tau)$ and $g_{it}(\tau)$ satisfy the following conditions : $f_{it}(\tau) \leq g_{it}(\tau)$, $f_{it}(\tau) > 0$, $g_{it}(\tau) > 0$.

I have designed two neural networks, one will be trained to compute $f_{it}$ and one will be trained to output $h_{it}$ where $g_{it} = f_{it} + h_{it}$. I impose non-negativity on outputs of both models such that the combined exit intensity will never be smaller than the default intensity for all horizons. Let us denote by $\lambda$ and $\mu$ the set of parameters (weights) tuned in the neural network for $f_{it}$ and $h_{it}$ respectively. $N^{(\lambda)}$ and $N^{(\mu)}$ stand for the output of the neural network for $f_{it}$ and $h_{it}$ respectively. The log-likelihood for prediction horizon $\tau$ is expressed\footnote{Proof of the above formulation can be found in the Appendix \ref{appendix1}.} as
\begin{equation} \label{decomp1}
\mathcal{L}(\lambda(s)) = \sum_{i=1}^N \sum_{t=0}^{T-s-1} \mathcal{L}_{i,t}(\lambda(s)), \qquad s = 0,1,...,\tau-1
\end{equation}

@@

As in previous studies, the likelihood functions still exhibit the decomposable property which allows to estimate the model for each horizon of prediction independently.

Since the intensities are directly driven by the covariates, \citet{Duan2012} require an assumption on the mapping from the covariates to these intensities. In \citet{Duan2012} the mapping is made with a linear assumption, whereas in the framework of this chapter, the mapping depends on the whole architecture of the neural network. When the neural network has only one hidden layer of one neuron coupled with an exponential activation function, the model boils down to \citet{Duan2012} as the intensities would be a linear combination of the covariates. As the width and depth of the network increases, we depart more and more from the linear assumption and we allow more non-linearities to be incorporated in these intensities. \\

@@

Neural networks can be seen as a very general function to map a given input (in this case a set of macroeconomic and firm-specific variables) into a desired output (forward intensities). It learns how to compute the output by tuning weights in order to minimize a given loss function. A neural network is constructed by juxtaposing several hidden layers of neurons. The input of each layer is a data transformation of the output of the previous layer. Initially, the weights of the network are assigned random values. Then, the training process starts and consists of many iterations of a forward pass and a backward pass. The forward pass takes as input a batch of data and computes the loss value, the backward pass then computes the gradient and adjust the weights of the network based on a learning rate hyperparameter. As an illustration, Figure \ref{fig:net} shows a neural network with 2 hidden layers : 5 neurons in the first layer and 3 neurons in the second layer. In this example, 3 features (inputs) are fed to the network. Each feature is connected to the first hidden layer by a set of weights. The outputs of the first layer are also weighted to produce the inputs of the second hidden layer. Non-linearity is introduced in each node by a non-linear activation function (e.g. sigmoid). Finally, the output of the second hidden layer is aggregated to produce the final output of the model. The neural networks in this chapter are implemented in Python using the library TensorFlow. Approximately one hour of computing time is needed to fit all networks on a 32GB RAM quad core 2.7 GHz computer.\footnote{GPU computing is not necessary in this context as the networks are small.}

@@

    % Annotate the layers
    \node[annot,left of=hl1] {Input Layer};
    \node[annot,right of=hl\Nhidden] {Output Layer};
\end{tikzpicture}
    \mycaption{Illustration of a neural network \text{[}5, 3\text{]}}{In this network, there are three input features (green circles), five neurons in the first hidden layer and 3 neurons in the second hidden layer. Each neuron is activated with a sigmoid function.}
    \label{fig:net}
\end{figure}
